As president of the European Association for Machine Translation (EAMT) and General Chair of the 25th Annual Conference of the EAMT, it is with great pleasure that I write these opening words to the Proceedings of EAMT 2024, a special year since we are celebrating our 25th anniversary!
\\

According to tradition, my first note of deep appreciation and gratitude goes to Celia Rico, Luc Meertens, Lucia Specia, and Maja Popovič, Executive Board Members, who have moved to new adventures in their lives, after outstanding, and dedicated service to the EAMT community.
\\

We have several milestones to celebrate this year, built upon the hard work of our Execu- tive Committee and our community: upgraded grants for low income and war zones and for Translation Studies, a record submission rate for research projects, continuous excelling submissions for the best thesis award, and one of the highest number of papers ever submitted to our conference (80 papers accepted)! I could not be prouder of our EC and the dynamics of our community.
\\

The EAMT Executive Committee (EC) has been very busy. Luc Meertens (treasurer), Carolina Scarton (secretary) and Sara Szoc (preparing to become our secretary and supporting everything we do) have been tirelessly supporting all initiatives. André Martins and Celia Rico, our co-chairs for low income areas, war zones and Translation Studies grants, selected 11 grantees, 6 applicants from Translation Studies and 5 from war zones (3 hybrid light and 8 in-person). Maja Popovič and Sara Szoc, our co-chairs for the Research Projects, selected 4  projects (equally distributed by students and general research projects calls) with a diverse set of topics. To all our co-chairs, my gratitude! The selection work is never an easy task and this year was particularly hard. 
\\

The same applied to the best thesis award –  Barry Haddow, chair of the Best Thesis Award, had a very difficult time selecting a candidate, since the submissions were of very high quality. Our congratulations to Marco Gaido's thesis “Direct Speech Translation Toward High-Quality, Inclusive, and Augmented Systems" (FBK, Italy), supervised by Marco Turchi and Matteo Negri. Our congratulations extended to the two highly commended theses of Jannis Vamvas: “Model-based Evaluation of Multilinguality” (University of Zurich, Switzerland), supervised by Rico Sennrich and Lena A. Jäger; and Javier Iranzo-Sánchez: “Streaming Neural Speech Translation” ( UPV, Spain), supervised by Jorge Civera and Alfons Juan. 
\\

EAMT, as full sponsor of the MT Marathon, would also like to highlight the outstanding work that the MT Marathon organisers conducted, enriching the vitality of our community with their projects and keynotes. A special thank you to the organising committee Lisa Yankovskaya, Agnes Luhtaru, Lisa Korotkova, Mark Fišel, Ondrej Bojar, and Barry Haddow for all the efforts on yet another successful MT Marathon event. Thank you, University of Tartu, for hosting the event.
\\

Sheffield, United Kingdom! EAMT 2024 celebrates our 25th anniversary! Our conference will have a three-day, four-track programme put together by our chairs: Rachel Bawden and Víctor Sánchez-Cartagena (research: technical track co-chairs); Ekaterina Lapshinova-Koltunski and Patrick Cadwell (research: translators \& users track co-chairs); Chatzitheodorou Konstantinos and Vera Cabarrão (implementations \& case studies track co-chairs); and Mikel Forcada and Helena Moniz (products \& projects track chairs). And backing up all the scientific components of our conference and filters of quality for the final selection: our reviewers. Thank you for your work and the alignment between all the chairs!
\\

Continuing the successful event from Tampere, this year EAMT 2024 will also have an extra day for workshops and tutorials, organised by our co-chairs Diptesh Kanojia and Mary Nurminen. Once more, the submissions for workshops and tutorials largely exceeded our expectations for our second edition!
\\

The programme will continue the tradition of including two keynote speakers, Alexandra Birch (Reader in Natural Language Processing in the Institute for Language, Cognition and Computation, School of Informatics, University of Edinburgh) and Valter Mavrič (Director-General of the Translation Service – DG TRAD – at the European Parliament). Our outstanding keynote speakers will demonstrate their extensive and global impactful work in translation studies and translation technologies.
\\

EAMT 2024 would never be possible without the synergetic, sharp, enthusiastic, and hard working local organising team! What a dream and fun team to work with! Our local co-chair, Carolina Scarton (University of Sheffield, UK), who always supports the EAMT community and is always eager to do the best EAMT ever! Our local co-chair from Zoo digital, Chris Oakley, also Charlotte Prescott (ZOO Digital, UK), Chris Bayliss (ZOO Digital, UK), Joanna Wright (University of Sheffield, UK), and Xingyi Song (University of Sheffield, UK). From the local organising support team, our thank you to Freddy Heppell (University of Sheffield, UK) and Tom Pickard (University of Sheffield, UK). Our special gratitude to the University of Sheffield and Zoo Digital for the joint efforts. You will surely make our 25th anniversary memorable!
\\

The Sheffield team is working towards a special 25th anniversary. Carolina Scarton has been doing intensive work on organising and finding a home for the John Hutchins Machine Translation Archive. Carolina is deeply committed to respect John’s wishes of making his library available to the community, and the former president, Mikel Forcada, and current one are fully supporting Carolina’s initiatives. As an anticipation of such effort, the Sheffield team is working on presenting a sample of John’s books for EAMT 2024 participants! Thank you, Carolina Scarton, for all the hard work on this. Within this topic still, a special thank you to Mike Hutchins, John’s son, who is fully committed to make it happen and respect his father’s vision of giving back to the community.
\\

EAMT has been supported by generous sponsors in its initiatives along the years. This year is no exception. Our gratitude to our Silver sponsors: RWS Language Weaver, Translated, and Unbabel. To our Bronze sponsors: CrossLang, Pangeanic, STAR, and TransPerfect. Also to Apertium, our long standing collaborator sponsor, Springer, our Supporter sponsor for the Best Paper award, and our Media sponsors, MultiLingual. Your support is vital in our efforts to give back to our community through grants and other initiatives.
\\

A note still to all our EAMT members and our participants! Without you no effort would make sense! Let us take this opportunity to create scientific collaboration and give constructive feedback. To fully enjoy the conference, please check our Code of Conduct at \url{https://eamt2024.sheffield.ac.uk/code-of-conduct}. I’m looking forward to seeing you all and celebrating our 25th anniversary with you!
\\

It is our organisation's greatest wish to continue giving back to our community and to drive and be driven by our community’s energy and enthusiasm. Reach out to us if you have new ideas or suggestions you would like to implement. We will try hard to accomplish it with you. Learn more about us at \url{https://eamt.org/}.

\bigbreak

Helena Moniz


\bigbreak


President of the EAMT

General Chair of EAMT 2024

University of Lisbon / INESC-ID, Portugal

